---
title: My dynamic report
output: html_fragment
---

```{r}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
```

```{r}
library(tidyverse)

#import Data from google sheets
logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vReF3vvFTYztjwdvP619aEGeKT4gSAO3CgVGqMzVhwYX-cG1owc7YRPli22YF7K8-uo9QDpu_4cFxUd/pub?gid=743329457&single=true&output=csv")


#Renamed columns of imported data 
latest_data <- logged_data %>% rename(Bus_Stop_Arrival_Time = 2,
  At_Mobile_Wait_Time = 3,
  Bus_Scheduled_Arrival_Time = 4,
  Departure_Bus_Stop_Number = 5, 
  Boariding_Bus_Number = 6,
  Day_Of_Week = 7,
  Time_Boarded_Bus = 8,
  AT_Mobile_Occupancy_Status= 9,
  Bus_Type = 10,
  Double_Empty_Seats_Available = 11,
  Sitting_Or_Standing = 12,
  Time_Off_Bus = 13,
  Destination_Bus_stop_number = 14,
  Weather = 15
)

#Set TimeStamp column date from chr to date 
latest_data$Date <- as.Date(
  latest_data$Timestamp,
  format = "%m/%d/%Y %H:%M:%S"
)

# Convert Bus_Stop_Arrival_Time data type chr to date-time 
latest_data$Bus_Stop_Arrival_TS <- as.POSIXct(
  paste(latest_data$Date, latest_data$Bus_Stop_Arrival_Time),
  format = "%Y-%m-%d %H:%M:%S"
)

# Convert Time_Boarded_Bus data type chr to date-time 
latest_data$Time_Boarded_Bus_TS <- as.POSIXct(
  paste(latest_data$Date, latest_data$Time_Boarded_Bus),
  format = "%Y-%m-%d %H:%M:%S"
)

# Convert Time_Off_Bus data type chr to date-time 
latest_data$Time_Off_Bus_TS <- as.POSIXct(
  paste(latest_data$Date, latest_data$Time_Off_Bus),
  format = "%Y-%m-%d %H:%M:%S"
)

# Generate new column Trip duration calculated  from Time_Off_Bus_TS - Time_Boarded_Bus_TS
latest_data$Trip_Duration <- latest_data$Time_Off_Bus_TS - latest_data$Time_Boarded_Bus_TS


# Generate new column actual_bus_wait_time from Time_Boarded_Bus_TS - Bus_Stop_Arrival_TS
latest_data$Actual_Wait_Time <- latest_data$Time_Boarded_Bus_TS - latest_data$Bus_Stop_Arrival_TS

#Convert Actual_Wait_Time from seconds to minutes 
latest_data$Actual_Wait_Minutes <- as.numeric(latest_data$Actual_Wait_Time, units = "mins")


```

```{css}

```

## Summary of Bus Wait Times

**Nature of the data:**  
This section uses **quantitative, time-based data**. The variables `Actual_Wait_Minutes` and `At_Mobile_Wait_Time` record numeric wait times in minutes for each bus trip. The data is observational and consists of repeated measurements collected at multiple timestamps for different trips.

**What I am analysing:**  
I am analysing differences between the **actual time waited at the bus stop** and the **predicted wait time from the AT Mobile app**. Calculating the **minimum, maximum, and average wait times** allows identification of trends in bus service reliability and the consistency of app predictions.


```{r}
# Summary for Actual_Wait_Minutes
paste("The minimum time I actually waited for a bus to come was:", 
      min(latest_data$Actual_Wait_Minutes, na.rm = TRUE), "minutes")

paste("The maximum time I actually waited for a bus to come was:", 
      max(latest_data$Actual_Wait_Minutes, na.rm = TRUE), "minutes")

paste("The average time I actually waited for a bus to come was:", 
      round(mean(latest_data$Actual_Wait_Minutes, na.rm = TRUE), 2), "minutes")

# Summary for At_Mobile_Wait_Time
paste("The minimum wait time for a bus to come recorded on At Mobile was:", 
      min(latest_data$At_Mobile_Wait_Time, na.rm = TRUE), "minutes")

paste("The maximum wait time for a bus to come recorded AT Mobile was:", 
      max(latest_data$At_Mobile_Wait_Time, na.rm = TRUE), "minutes")

paste("The average wait time for a bus to come recorded At Mobile was:", 
      round(mean(latest_data$At_Mobile_Wait_Time, na.rm = TRUE), 2), "minutes")

```

## Bar charts

## Actual_Wait_Minutes vs AT_Mobile_Wait_Time Bar Chart

**Nature of the data:**  
This plot visualises **quantitative time-based variables** (`Actual_Wait_Minutes` and `AT_Mobile_Wait_Time`) for each trip. Each timestamp represents a repeated observation of a single bus journey.

**What I am analysing:**  
The chart compares **actual vs predicted wait times** to determine whether buses arrive earlier or later than expected, and to evaluate the accuracy and consistency of the app’s predictions over multiple trips.


```{r}


# Actual_Wait_Minutes vs At_Mobile_Wait_Time
ggplot(latest_data, aes(x = Timestamp)) +
  geom_col(aes(y = Actual_Wait_Minutes), fill = "blue", width = 0.4) +
  geom_col(aes(y = At_Mobile_Wait_Time), fill = "red", width = 0.4, position = position_nudge(x = 0.2)) +
  labs(title = "Actual Wait Minutes vs AT Mobile Wait Time",
       subtitle =  "blue = Actual Wait Minutes, red = AT Mobile Wait Time",
       caption = "Source: My Obsevation log",
       x = "Date", 
       y = "Wait Time (minutes)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

## Bus Occupancy by Day of the Week

**Nature of the data:**  
This section uses **categorical data**. `AT_Mobile_Occupancy_Status` indicates crowding levels on the bus, while `Day_Of_Week` identifies the day each observation was made. The data is observational and collected over multiple trips.

**What I am analysing:**  
I am analysing how **bus occupancy varies by day of the week**, identifying patterns in crowding. This shows which days tend to have more standing passengers or more available seats, highlighting trends in passenger load across the week.

```{r}
#AT_Mobile_Occupancy_Status vs Day_Of_Week
latest_data %>%
  ggplot() +
  geom_bar(aes(y = AT_Mobile_Occupancy_Status, 
               fill = Day_Of_Week)) +
  labs(title = "AT Mobile Occupancy Status vs Weather",
       caption = "Source: My Obsevation log",
       x = "AT Mobile Occupancy Status Count",
       y = "AT Mobile Occupancy Status type ")
```

## Bus Occupancy by Weather

**Nature of the data:**  
This section also uses **categorical data**. `AT_Mobile_Occupancy_Status` shows crowding levels, while `Weather` records conditions during each trip. Data is repeated over multiple trips and weather types.

**What I am analysing:**  
I am analysing how **bus occupancy changes under different weather conditions**, to understand whether weather affects crowding and passenger comfort. The plot helps reveal trends such as more standing passengers on rainy days or more available seating on sunny days.

```{r}
#AT_Mobile_Occupancy_Status vs Weather",
latest_data %>%
  ggplot() +
  geom_bar(aes(y = AT_Mobile_Occupancy_Status, fill = Weather)) +
  labs(title = "AT Mobile Occupancy Status vs Weather",
       caption = "Source: My Obsevation log",
       x = "AT Mobile Occupancy Status Count",
       y = "AT Mobile Occupancy Status type ")

```
