---
title: My dynamic report
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

library(tidyverse)
options(scipen=999)

```

```{css}
.dynamic-comment {
background-color: lightblue;
}
```

```{r}
form_url <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vT5NlRQI6v0Rn42uRLOu3cGidKE5gROx3toExiqMiIm4XUpytKAw1lxzhMxvYepOVOvoyt-pSCbH7g_/pub?gid=1028618258&single=true&output=csv"

survey_data <- read_csv(form_url) 

survey_data_renamed <- survey_data %>%
  rename(prev_course = 2,
         programming_confidence = 3,
         parttime_job = 4,
         parttime_hours = 5,
         pizza_pref = 6,
         height_cm = 7,
         adopt_pet = 8,
         positive_thing = 9,
         upi_example = 10)
```

## Part-time work during semester?

```{r, echo= F}
survey_data_renamed %>% 
  ggplot(aes(x = parttime_job)) +
  geom_bar() +
  theme_minimal() +
  labs(title = "Part-time work", subtitle = "STATS 220, 2026",
       x = "Do you currently have a part-time job during the university semester?")
```

# Part-time hours

```{r, echo = F}
hours_mean <- survey_data_renamed$parttime_hours %>% 
  mean(na.rm = TRUE) 
```

::: dynamic-comment
The median reported time was `r median(survey_data_renamed$parttime_hours, na.rm = TRUE)`. For students who have a part-time job, the average number of hours they work per week is `r round(hours_mean, 1)`. The highest reported hour amount was `r max(survey_data_renamed$parttime_hours, na.rm = TRUE)`. Ooops, I didn't add sensible data validation to my survey. I hope I will learn how to filter and clean my data in the future!
:::
